---
title: "data_download"
author: "Jan Besler"
date: "`r Sys.Date()`"
output: html_document
---
### setup

```{r setup, include=FALSE}
# load necessary libraries
library(lubridate)
library(readr)

knitr::opts_chunk$set(echo = TRUE)
```

# stock and certificate prices

download the needed historical data from [yahoo finance](https://finance.yahoo.com/) and [onvista](https://www.onvista.de/derivate/Express-Zertifikate/handelsplaetze/204517484-UBS1KS-DE000UBS1KS0)

> The company is not paying a dividend to its shareholders a stated in the latest financial report for 2021.
*https://www.morphosys.com/sites/reports/en/gb/2021/*

```{r download}
# download newest stock data
MOR <-
  read_csv(url(
    paste0(
      "https://query1.finance.yahoo.com/v7/finance/download/MOR.DE?period1=1524096000&period2=",
      as.integer(as.POSIXct(Sys.time())),
      "&interval=1d&events=history&includeAdjustedClose=true"
    )
  ), lazy = FALSE)

# load the api keys
source("./data/apikeys.R")
# read in the downloaded csv file
ExZ <- read_csv2("./data/ExZ_Data_latest.csv", col_names = TRUE)
  

# remove duplicates
MOR <- MOR[!duplicated(MOR$Date), ]
```

### tidy data

tidy the data to make the columns compatible later on.

```{r tidying}
sapply(MOR, class)
sapply(ExZ, class)

# change 'Datum' type since 'Date' and 'Datum' are not the same type
ExZ$Datum <- as.Date(parse_date_time(ExZ$Datum, c('dmy', 'ymd')))
sapply(ExZ, class)
```

### save data

store the data as csv files in the data folder

```{r store}
write_csv(MOR, "./data/MOR.csv", col_names = TRUE)

write_csv(ExZ, "./data/ExZ_Data.csv", col_names = TRUE)
```

# download the svensson parameters

The data is downloaded from [Bundesbank](https://www.bundesbank.de/dynamic/action/en/statistics/time-series-databases/time-series-databases/759784/759784?listId=www_skms_it03c)
The data starts at the 01.08.1997
```{r svensson, warning=FALSE}
startDate <- "1997-08-01"

# list all required parameters
svenssonParameters <- c("B0", "B1", "B2", "B3", "T1", "T2")

# data frame dimensions
i = 13
j = length(seq(as.Date(startDate), as.Date(Sys.time()), by = 'day')) - 2
  
# build empty data frame
svensson <- data.frame(matrix(NA, nrow = j, ncol = i))

# loop to download all necessary parameters and store in a single data frame
for(i in 1:length(svenssonParameters)){
svenssonRaw <- read_csv(url(
  paste0("https://api.statistiken.bundesbank.de/rest/download/BBSIS/D.I.ZST.",
  svenssonParameters[i],
  ".EUR.S1311.B.A604._Z.R.A.A._Z._Z.A?format=csv&lang=en")),
  lazy = FALSE,
  na = c(".", "No value available", "Kein Wert vorhanden", "NA", NA),
  skip = 8,
  col_names = FALSE)
# store in data frame
svensson[,(i*2):(i*2+1)] <- svenssonRaw[,1:2]
}

# tidying the data set
svensson <- svensson[,-c(1, 4, 6, 8, 10, 12)]
colnames(svensson)[c(1:7)] <- c("Date", "b0", "b1", "b2", "b3", "t1", "t2")
svensson <- svensson[!(is.na(svensson$b0)),]

# save the table as csv
write_csv(svensson, "./data/svensson.csv", col_names = TRUE)
```

```{r stock_rm}
rm(
  ExZ,
  MOR,
  svensson,
  svenssonParameters,
  svenssonRaw
)
```