---
title: "greeks"
author: "Jan Besler"
date: "`r Sys.Date()`"
output: html_document
---

# Setup

> ***libraries***

before starting the work space is cleared and all necessary libraries are loaded.

```{r setup, include=FALSE}
# load libraries
library(knitr)      # use .Rmd files

  # reading data
library(readr)      # reading csv

  # manipulate data
library(tidyr)      # manipulating data
library(dplyr)      # manipulating data
library(rlist)      # working with lists
library(stringr)    # work with strings
library(tibble)     # create data frames
library(stringi)    # process charactar strings
library(lubridate)  # transform character into dates

  # statistics
library(Matrix)     # required for lme4

  # visuals
library(ggplot2)    # make graphs
library(ggExtra)    # additional functionality for ggplot2

knitr::opts_chunk$set(echo = TRUE)
```

# Part V

For the ***Sensitivity Analysis*** three timestamps before the next valuation date are chosen. These timestamps are six, three and one month in advance of, in this case, *24.08.2022*.
The Greeks are estimated just as before using a simulation with `N = 1000`.

> The last date is changed to match with the latest data available, since original day is the day of handing in the modul

```{r V_setup}
# get indices for corresponding dates
greekDates <- as.Date(c("2022-02-24", "2022-05-24", "2022-07-15"))
  
# define price range to use
StockPriceRange <- seq(15, 55, 0.2)
```

Since the Greeks are different derivatives of the Black Scholes formula and this can not be implemented directly. The derivatives have to approximated by using numerical differentiation, in particular the method of *finite differences* has been chosen.

### Delta

##### Function

To compute the finite differences for delta, the Certificate Price is adjusted by `+- 0.5`

```{r delta}
# define delta function
Delta <- function(startPrice, riskFreeDaily, vola, T0, N){
  # determine magnitude of change
  finiteChange <- 0.5
  # same computation used for Certificate Pricing as before
    # store values used as index in vector
  pathsResult <- matrix(data = NA, nrow = length(daysToMaturity), ncol = N)
  CertPlus <- vector(mode = "numeric", length = length(StockPriceRange))
  CertMinus <- vector(mode = "numeric", length = length(StockPriceRange))
  
# StockPrice + Change
  StockPrice <- startPrice + finiteChange
    # same simulation as used previously
  simulatedPaths <- simulation(StockPrice, riskFreeDaily, vola, T0, N)
  # derive payoffs by looking at observation dates
    # reverse matrix for easier indexing
  simulatedInverse <- apply(simulatedPaths, 2, rev)
  # extract rows for evaluation dates
    # take number to maturity as index
  for (k in 1:length(daysToMaturity)){
    # for every valuation date possible store row in extra matrix
    if (daysToMaturity[k] < nrow(simulatedInverse))
      pathsResult[k,] <- simulatedInverse[daysToMaturity[k],]
    # otherwise put 0 in to fail at payoff function
    else
      pathsResult[k,] <- 0
  }
  # build certificate price as mean discounted payoff from stock prices
  CertPlus[j] <-
    as.numeric(exp(-riskFreeDaily * (T0/240)) * mean(payoff(pathsResult), na.rm = TRUE))

# StockPrice - Change
  StockPrice <- startPrice - finiteChange
    # same simulation as used previously
  simulatedPaths <- simulation(StockPrice, riskFreeDaily, vola, T0, N)
  # derive payoffs by looking at observation dates
    # reverse matrix for easier indexing
  simulatedInverse <- apply(simulatedPaths, 2, rev)
  # extract rows for evaluation dates
    # take number to maturity as index
  for (k in 1:length(daysToMaturity)){
    # for every valuation date possible store row in extra matrix
    if (daysToMaturity[k] < nrow(simulatedInverse))
      pathsResult[k,] <- simulatedInverse[daysToMaturity[k],]
    # otherwise put 0 in to fail at payoff function
    else
      pathsResult[k,] <- 0
  }
  # build certificate price as mean discounted payoff from stock prices
  CertMinus[j] <-
    as.numeric(exp(-riskFreeDaily * (T0/240)) * mean(payoff(pathsResult), na.rm = TRUE))
  # return results
  return(CertPlus[j] - CertMinus[j])
}
```

##### Calculation

Calculate Delta for each of the pre defined dates and each stock values defined in the Stockprice range.

```{r Delta2}
# create matrix to store values in
DeltaMatrix <- matrix(
  data = NA,
  nrow = length(StockPriceRange),
  ncol = length(greekDates)
)
  # helper matrices
DeltaPlusMatrix <- matrix(
  data = NA,
  nrow = length(StockPriceRange),
  ncol = length(greekDates)
)
DeltaMinusMatrix <- matrix(
  data = NA,
  nrow = length(StockPriceRange),
  ncol = length(greekDates)
)

GreekData <- as.data.frame(matrix(data = NA, ncol = 10, nrow = length(greekDates)))
colnames(GreekData) <- names(dataSim)
  


# extract rows for specified dates
for (i in 1:length(greekDates)) {
  GreekData[i, ] <-
    dataSim[dataSim$Date == greekDates[i], ]
  # define variables for simulation
  riskFreeDaily <- GreekData$riskFreeDaily[[i]]
  vola <-
    GreekData$volaY2[[i]] / sqrt(240)  # use volatility scaled to daily value
  T0 <- GreekData$untilmaturity[[i]]
  # calculate for using the respective variables and StockPriceRange for startPrice and same N as before
  for (j in 1:length(StockPriceRange)) {
    DeltaMatrix[j, i] <- Delta(StockPriceRange[j],
                                 riskFreeDaily,
                                 vola,
                                 T0,
                                 N)
  }
}
```

##### Visualization

Plot the different Delta against the stock price.

```{r Delta3, warning=FALSE}
# add stockprice column to matrix
DeltaMatrix <- as.data.frame(cbind(DeltaMatrix, StockPriceRange))

# plot Delta
deltaplot <- ggplot(
  data = DeltaMatrix,
  aes(x = StockPriceRange)) +
  geom_line(aes(y = V1, color = '6 months before observation date')) +
  geom_line(aes(y = V2, color = '3 months before observation date')) +
  geom_line(aes(y = V3, color = '~1 month before observation date')) +
  geom_vline(xintercept = c(initialStockPrice, initialStockPrice * 0.6), linetype = 'dashed', color = 'black') +
  xlim(25, 55) +
  labs(x = "Stock Price", y = "") +
  ggtitle("Delta",
          "with Barrier and Strike Price") +
  theme_bw(base_size = 15) +
  theme(legend.position = c(0.75, 0.8),
        legend.title = element_blank())

# save plot 
ggsave(
  filename = "./figures/DeltaPlot.png",
  dpi = 320
)
```

### Gamma

##### Function

For the calculation of Gamma, the function `CertificatePrice + 0.5 - 2 * CertificatePrice + CertficatePrice - 0.5` is used.

```{r gamma1}
# define delta function
Gamma <- function(startPrice, riskFreeDaily, vola, T0, N){
  # determine magnitude of change
  finiteChange <- 0.5
  # same computation used for Certificate Pricing as before
    # store values used as index in vector
  pathsResult <- matrix(data = NA, nrow = length(daysToMaturity), ncol = N)
  Cert <- vector(mode = "numeric", length = length(StockPriceRange))
  CertPlus <- vector(mode = "numeric", length = length(StockPriceRange))
  CertMinus <- vector(mode = "numeric", length = length(StockPriceRange))
  
# StockPrice
  StockPrice <- startPrice
    # same simulation as used previously
  simulatedPaths <- simulation(StockPrice, riskFreeDaily, vola, T0, N)
  # derive payoffs by looking at observation dates
    # reverse matrix for easier indexing
  simulatedInverse <- apply(simulatedPaths, 2, rev)
  # extract rows for evaluation dates
    # take number to maturity as index
  for (k in 1:length(daysToMaturity)){
    # for every valuation date possible store row in extra matrix
    if (daysToMaturity[k] < nrow(simulatedInverse))
      pathsResult[k,] <- simulatedInverse[daysToMaturity[k],]
    # otherwise put 0 in to fail at payoff function
    else
      pathsResult[k,] <- 0
  }
  # build certificate price as mean discounted payoff from stock prices
  Cert[j] <-
    as.numeric(exp(-riskFreeDaily * (T0/240)) * mean(payoff(pathsResult), na.rm = TRUE))

  
# StockPrice + Change
  StockPrice <- startPrice + finiteChange
    # same simulation as used previously
  simulatedPaths <- simulation(StockPrice, riskFreeDaily, vola, T0, N)
  # derive payoffs by looking at observation dates
    # reverse matrix for easier indexing
  simulatedInverse <- apply(simulatedPaths, 2, rev)
  # extract rows for evaluation dates
    # take number to maturity as index
  for (k in 1:length(daysToMaturity)){
    # for every valuation date possible store row in extra matrix
    if (daysToMaturity[k] < nrow(simulatedInverse))
      pathsResult[k,] <- simulatedInverse[daysToMaturity[k],]
    # otherwise put 0 in to fail at payoff function
    else
      pathsResult[k,] <- 0
  }
  # build certificate price as mean discounted payoff from stock prices
  CertPlus[j] <-
    as.numeric(exp(-riskFreeDaily * (T0/240)) * mean(payoff(pathsResult), na.rm = TRUE))

# StockPrice - Change
  StockPrice <- startPrice - finiteChange
    # same simulation as used previously
  simulatedPaths <- simulation(StockPrice, riskFreeDaily, vola, T0, N)
  # derive payoffs by looking at observation dates
    # reverse matrix for easier indexing
  simulatedInverse <- apply(simulatedPaths, 2, rev)
  # extract rows for evaluation dates
    # take number to maturity as index
  for (k in 1:length(daysToMaturity)){
    # for every valuation date possible store row in extra matrix
    if (daysToMaturity[k] < nrow(simulatedInverse))
      pathsResult[k,] <- simulatedInverse[daysToMaturity[k],]
    # otherwise put 0 in to fail at payoff function
    else
      pathsResult[k,] <- 0
  }
  # build certificate price as mean discounted payoff from stock prices
  CertMinus[j] <-
    as.numeric(exp(-riskFreeDaily * (T0/240)) * mean(payoff(pathsResult), na.rm = TRUE))
  # return results
  return(CertPlus[j] + 2*Cert[j] - CertMinus[j])
}
```

##### Calculation

```{r gamma2}
# create matrix to store values in
GammaMatrix <- matrix(
  data = NA,
  nrow = length(StockPriceRange),
  ncol = length(greekDates)
)

# extract rows for specified dates
for (i in 1:length(greekDates)) {
  GreekData[i, ] <-
    dataSim[dataSim$Date == greekDates[i], ]
  # define variables for simulation
  riskFreeDaily <- GreekData$riskFreeDaily[[i]]
  vola <-
    GreekData$volaY2[[i]] / sqrt(240)  # daily volatility value
  T0 <- GreekData$untilmaturity[[i]]
  # calculate for using the respective variables and StockPriceRange for startPrice and same N as before
  for (j in 1:length(StockPriceRange)) {
    GammaMatrix[j, i] <- Gamma(StockPriceRange[j],
                                 riskFreeDaily,
                                 vola,
                                 T0,
                                 N)
  }
}
```

##### Visualization

Plot the different Gamma against the stock price.

```{r Delta3, warning=FALSE}
# add stockprice column to matrix
GammaMatrix <- as.data.frame(cbind(GammaMatrix, StockPriceRange))

# plot Delta
gammaplot <- ggplot(
  data = GammaMatrix,
  aes(x = StockPriceRange)) +
  geom_line(aes(y = V1, color = '6 months before observation date')) +
  geom_line(aes(y = V2, color = '3 months before observation date')) +
  geom_line(aes(y = V3, color = '~1 month before observation date')) +
  geom_vline(xintercept = c(initialStockPrice, initialStockPrice * 0.6), linetype = 'dashed', color = 'black') +
  xlim(25, 55) +
  labs(x = "Stock Price", y = "") +
  ggtitle("Gamma",
          "with Barrier and Strike Price") +
  theme_bw(base_size = 15) +
  theme(legend.position = c(0.75, 0.2),
        legend.title = element_blank())

# save plot 
ggsave(
  filename = "./figures/GammaPlot.png",
  dpi = 320
)
```

### Vega

##### Function

For the calculation of Vega, the is volatility measures are changed.
After trying different values the values *5%* and *45%* have been chosen.

```{r vega1}
# define delta function
Vega <- function(startPrice, riskFreeDaily, vola, T0, N){
  # determine magnitude of change
  finiteChange <- 0.5
  # same computation used for Certificate Pricing as before
    # store values used as index in vector
  pathsResult <- matrix(data = NA, nrow = length(daysToMaturity), ncol = N)
  VolaLow <- vector(mode = "numeric", length = length(StockPriceRange))
  VolaHigh <- vector(mode = "numeric", length = length(StockPriceRange))
  
# low volatility
  StockPrice <- startPrice
  volalow <- 0.05
    # same simulation as used previously
  simulatedPaths <- simulation(StockPrice, riskFreeDaily, volalow, T0, N)
  # derive payoffs by looking at observation dates
    # reverse matrix for easier indexing
  simulatedInverse <- apply(simulatedPaths, 2, rev)
  # extract rows for evaluation dates
    # take number to maturity as index
  for (k in 1:length(daysToMaturity)){
    # for every valuation date possible store row in extra matrix
    if (daysToMaturity[k] < nrow(simulatedInverse))
      pathsResult[k,] <- simulatedInverse[daysToMaturity[k],]
    # otherwise put 0 in to fail at payoff function
    else
      pathsResult[k,] <- 0
  }
  # build certificate price as mean discounted payoff from stock prices
  VolaLow[j] <-
    as.numeric(exp(-riskFreeDaily * (T0/240)) * mean(payoff(pathsResult), na.rm = TRUE))

# high volatility
  StockPrice <- startPrice
  volahigh <- 0.45
    # same simulation as used previously
  simulatedPaths <- simulation(StockPrice, riskFreeDaily, volahigh, T0, N)
  # derive payoffs by looking at observation dates
    # reverse matrix for easier indexing
  simulatedInverse <- apply(simulatedPaths, 2, rev)
  # extract rows for evaluation dates
    # take number to maturity as index
  for (k in 1:length(daysToMaturity)){
    # for every valuation date possible store row in extra matrix
    if (daysToMaturity[k] < nrow(simulatedInverse))
      pathsResult[k,] <- simulatedInverse[daysToMaturity[k],]
    # otherwise put 0 in to fail at payoff function
    else
      pathsResult[k,] <- 0
  }
  # build certificate price as mean discounted payoff from stock prices
  VolaHigh[j] <-
    as.numeric(exp(-riskFreeDaily * (T0/240)) * mean(payoff(pathsResult), na.rm = TRUE))
  # return results
  return((VolaLow[j] - VolaHigh[j]) / (volahigh - volalow))
}
```

##### Calculation

```{r vega2}
# create matrix to store values in
VegaMatrix <- matrix(
  data = NA,
  nrow = length(StockPriceRange),
  ncol = length(greekDates)
)

# extract rows for specified dates
for (i in 1:length(greekDates)) {
  GreekData[i, ] <-
    dataSim[dataSim$Date == greekDates[i], ]
  # define variables for simulation
  riskFreeDaily <- GreekData$riskFreeDaily[[i]]
  vola <-
    GreekData$volaY2[[i]] / sqrt(240)  # daily volatility value
  T0 <- GreekData$untilmaturity[[i]]
  # calculate for using the respective variables and StockPriceRange for startPrice and same N as before
  for (j in 1:length(StockPriceRange)) {
    VegaMatrix[j, i] <- Vega(StockPriceRange[j],
                                 riskFreeDaily,
                                 vola,
                                 T0,
                                 N)
  }
}
```

##### Visualization

Plot the different Gamma against the stock price.

```{r vega3, warning=FALSE}
# add stockprice column to matrix
VegaMatrix <- as.data.frame(cbind(VegaMatrix, StockPriceRange))

# plot Vega
vegaplot <- ggplot(
  data = VegaMatrix,
  aes(x = StockPriceRange)) +
  geom_line(aes(y = V1, color = '6 months before observation date')) +
  geom_line(aes(y = V2, color = '3 months before observation date')) +
  geom_line(aes(y = V3, color = '~1 month before observation date')) +
  geom_vline(xintercept = c(initialStockPrice, initialStockPrice * 0.6), linetype = 'dashed', color = 'black') +
  xlim(25, 55) +
  labs(x = "Stock Price", y = "") +
  ggtitle("Vega",
          "with Barrier and Strike Price") +
  theme_bw(base_size = 15) +
  theme(legend.position = c(0.5, 0.2),
        legend.title = element_blank())

# save plot 
ggsave(
  filename = "./figures/VegaPlot.png",
  dpi = 320
)
```


### Theta

##### Function

For the calculation of Theta, the Certificate prices are compared to the prices one day closer to the expiration date.

```{r theta1}
# define delta function
Theta <- function(startPrice, riskFreeDaily, vola, T0, N){
  # same computation used for Certificate Pricing as before
    # store values used as index in vector
  pathsResult <- matrix(data = NA, nrow = length(daysToMaturity), ncol = N)
  CertCurrent <- vector(mode = "numeric", length = length(StockPriceRange))
  CertLagged <- vector(mode = "numeric", length = length(StockPriceRange))
  
# normal days to maturity
  StockPrice <- startPrice
  
    # same simulation as used previously
  simulatedPaths <- simulation(StockPrice, riskFreeDaily, vola, T0, N)
  # derive payoffs by looking at observation dates
    # reverse matrix for easier indexing
  simulatedInverse <- apply(simulatedPaths, 2, rev)
  # extract rows for evaluation dates
    # take number to maturity as index
  for (k in 1:length(daysToMaturity)){
    # for every valuation date possible store row in extra matrix
    if (daysToMaturity[k] < nrow(simulatedInverse))
      pathsResult[k,] <- simulatedInverse[daysToMaturity[k],]
    # otherwise put 0 in to fail at payoff function
    else
      pathsResult[k,] <- 0
  }
  # build certificate price as mean discounted payoff from stock prices
  CertCurrent[j] <-
    as.numeric(exp(-riskFreeDaily * (T0/240)) * mean(payoff(pathsResult), na.rm = TRUE))

# lagged days to maturity
  StockPrice <- startPrice
  T1 <- T0 - 1
    # same simulation as used previously
  simulatedPaths <- simulation(StockPrice, riskFreeDaily, vola, T1, N)
  # derive payoffs by looking at observation dates
    # reverse matrix for easier indexing
  simulatedInverse <- apply(simulatedPaths, 2, rev)
  # extract rows for evaluation dates
    # take number to maturity as index
  for (k in 1:length(daysToMaturity)){
    # for every valuation date possible store row in extra matrix
    if (daysToMaturity[k] < nrow(simulatedInverse))
      pathsResult[k,] <- simulatedInverse[daysToMaturity[k],]
    # otherwise put 0 in to fail at payoff function
    else
      pathsResult[k,] <- 0
  }
  # build certificate price as mean discounted payoff from stock prices
  CertLagged[j] <-
    as.numeric(exp(-riskFreeDaily * (T0/240)) * mean(payoff(pathsResult), na.rm = TRUE))
  # return results
  return((CertCurrent[j] - CertLagged[j]) / (T1 - T0))
}
```

##### Calculation

```{r theta2}
# create matrix to store values in
ThetaMatrix <- matrix(
  data = NA,
  nrow = length(StockPriceRange),
  ncol = length(greekDates)
)

# extract rows for specified dates
for (i in 1:length(greekDates)) {
  GreekData[i, ] <-
    dataSim[dataSim$Date == greekDates[i], ]
  # define variables for simulation
  riskFreeDaily <- GreekData$riskFreeDaily[[i]]
  vola <- GreekData$volaY2[[i]] / sqrt(240)  # daily volatility value
  T0 <- GreekData$untilmaturity[[i]]
  # calculate for using the respective variables and StockPriceRange for startPrice and same N as before
  for (j in 1:length(StockPriceRange)) {
    ThetaMatrix[j, i] <- Theta(StockPriceRange[j],
                                 riskFreeDaily,
                                 vola,
                                 T0,
                                 N)
  }
}
```

##### Visualization

Plot the different Theta against the stock price.

```{r theta3, warning=FALSE}
# add stockprice column to matrix
ThetaMatrix <- as.data.frame(cbind(ThetaMatrix, StockPriceRange))

# plot Theta
thetaplot <- ggplot(
  data = ThetaMatrix,
  aes(x = StockPriceRange)) +
  geom_line(aes(y = V1, color = '6 months before observation date')) +
  geom_line(aes(y = V2, color = '3 months before observation date')) +
  geom_line(aes(y = V3, color = '~1 month before observation date')) +
  geom_vline(xintercept = c(initialStockPrice, initialStockPrice * 0.6), linetype = 'dashed', color = 'black') +
  xlim(25, 55) +
  labs(x = "Stock Price", y = "") +
  ggtitle("Theta",
          "with Barrier and Strike Price") +
  theme_bw(base_size = 15) +
  theme(legend.position = c(0.75, 0.8),
        legend.title = element_blank())

# save plot 
ggsave(
  filename = "./figures/ThetaPlot.png",
  dpi = 320
)

ggarrange(deltaplot, gammaplot,
          vegaplot, thetaplot,
          ncol = 2,
          nrow = 2)
ggsave("./figures/Greeks.png",
       dpi= 320,
       width = 14,
       height = 7)

```
