---
title: "Parameter selection & valuation"
author: "Jan Besler"
date: "`r Sys.Date()`"
output: html_document
---

# Setup

> ***libraries***

before starting the work space is cleared and all necessary libraries are loaded.

```{r setup, include=FALSE}
# load libraries
library(knitr)      # use .Rmd files

  # reading data
library(readr)      # reading csv

  # manipulate data
library(tidyr)      # manipulating data
library(dplyr)      # manipulating data
library(rlist)      # working with lists
library(stringr)    # work with strings
library(tibble)     # create data frames
library(stringi)    # process charactar strings
library(lubridate)  # transform character into dates

  # statistics
library(Matrix)     # required for lme4

  # visuals
library(ggplot2)    # make graphs
library(ggExtra)    # additional functionality for ggplot2

knitr::opts_chunk$set(echo = TRUE)
```

# real data

inspect the data frames before working with them

```{r preperation}
data <- read_csv("./data/MOR.csv",
                col_select = c("Date", "Stock"),
                col_names = c("Date", "Stock"),
                skip = 1)
ExZ <- read_csv("./data/ExZ_Data.csv",
                col_select = c("Date", "Certificate"),
                col_names = c("Date", "Certificate"),
                skip = 1)

dim(data)
head(data, 5)

dim(ExZ)
head(ExZ, 5)

# look at data and classes
glimpse(data)
```

# Volatility

The volatility is calculated as the rolling standard deviation of the stock return.

```{r vola}
# define the number of trading Days for one and two years
traidingDays <- c(240, 480)

# create empty vectors to store the values
vola1 <- vector(mode = "numeric", length = length(data$Stock))
vola2 <- vector(mode = "numeric", length = length(data$Stock))
diff <- vector(mode = "numeric", length = length(data$Stock))

# calculate the volatility for 1 year
for(i in 2:length(data$Stock)){
  diff[i-1] <- log(data$Stock[[i-1]]) - log(data$Stock[[i]])  # daily stock return
  vola1[i-1] <- sd(diff) * sqrt(traidingDays[1])  # standard deviation of stock return for year 1
  vola2[i-1] <- sd(diff) * sqrt(traidingDays[2])  # same as before for year 2
}

# store data in new column
data$volaY1 <- vola1
data$volaY2 <- vola2

head(data, 5)
```

### days until maturity

first calculate the maximum number of days this certificate can be traded.
```{r trading days}
emission <- as.Date(min(data$Date)) # first date for which data of certificate exists
maturity <- as.Date("2026-08-31")    # taken from information sheet
sum(!weekdays(seq(emission, maturity, "days"))
    %in% c("Saturday", "Sunday"))
```

```{r maturitiy}
# create empty vector
durationExZ <- vector(mode = "integer", length = nrow(data))

# loop to calculate the remaining duration of the certificate
for(i in 1:length(durationExZ)){
durationExZ[i] <- sum(!weekdays(seq(as.Date(data[[1]][i]), maturity, "days"))
                    %in% c("Saturday", "Sunday"))/traidingDays[1]
}

# add vector as column into data frame
data$timeRemaining <- durationExZ

# define valuation dates
valuation <- as.Date(c("2022-08-24", "2023-08-24", "2024-08-26", "2025-08-25", "2026-08-31"))

# create vector 
nextValuation <- data.frame(matrix(NA, nrow = length(data$Stock), ncol = length(valuation)))

for(i in 1:length(durationExZ)){
  for(j in 1:length(nextValuation)){
    nextValuation[i,j] <- sum(!weekdays(seq(as.Date(data[[1]][i]), valuation[j], "days"))
                    %in% c("Saturday", "Sunday"))
  }
}

nextValuation <- nextValuation %>% dplyr::rename(
  until22 = X1,
  until23 = X2,
  until24 = X3,
  until25 = X4,
  untilmaturity = X5
)

# join the two data sets
nextValuation$Date <- data$Date
data <- inner_join(
  data,
  nextValuation,
  by = "Date"
)
glimpse(data)
```

### Risk free rate

we use the already prepared data set from Bundesbank.
```{r risk free}
# load data set
svensson <- read_csv("./data/svensson.csv")

# join the data
data <- inner_join(
  data,
  svensson,
  by = "Date"
)
glimpse(data)
```

The risk free rate and the daily risk free rate are computed using the [Svensson parameters](https://www.nber.org/papers/w4871).

```{r risk free, warning=FALSE}
# create empty vector to fill and attach to data frame
b1Exp <- vector(mode = "numeric", length = length(data$Stock))
b2Exp <- vector(mode = "numeric", length = length(data$Stock))
b3Exp <- vector(mode = "numeric", length = length(data$Stock))

# calculating the risk free rate using the Svensson functional relation
for(i in 1:nrow(data)){
    # define the exponential components used in risk free rate calculations
    b1Exp[i] <- (1 - exp(-data[['timeRemaining']][i]/data[['t1']][i])) / (data[['timeRemaining']][i]/data[['t1']][i])
    b2Exp[i] <- ((1 - exp(-data[['timeRemaining']][i]/data[['t1']][i])) / (data[['timeRemaining']][i]/data[['t1']][i])) - exp(-(data[['timeRemaining']][i]/data[['t1']][i]))
    b3Exp[i] <- ((1 - exp(-data[['timeRemaining']][i]/data[['t2']][i])) / (data[['timeRemaining']][i]/data[['t2']][i])) - exp(-(data[['timeRemaining']][i]/data[['t2']][i]))
    # put whole function together & store values in data frame
    data$riskFree[i] <- (data[['b0']][i]
            + data[['b1']][i]*b1Exp 
            + data[['b2']][i]*b2Exp 
            + data[['b3']][i]*b3Exp) / 100
}
# check for missing values
sum(is.na(data$riskFree))
# fill missing value up with previous value
data <- data %>% fill(riskFree)

# use the risk free rate to calculate the daily rate
data$riskFreeDaily <- data$riskFree / sqrt(traidingDays[1])

head(data, 5)

# save data to use later in Part 2
write_csv(data, "./data/dataPart2.csv", col_names = TRUE)
# join Certificate Prices
  # Only the days during which both products exist, can be used.
data <- inner_join(
  data,
  ExZ,
  by = "Date"
)

# remove unneeded data and variables
rm(
  svensson,
  b1Exp,
  b2Exp,
  b3Exp,
  nextValuation,
  durationExZ,
  vola1,
  vola2,
  ExZ,
  diff
)
```

# valuation functions

  1. function that defines payoff based on breaking the maturity barrier at evaluation dates
    - store payoff in vector 'payoff'
    - define payoff at maturity date
  2. brownian motion simulation n-times for all dates until maturity.
  3. definition of RSME for evaluation
  
```{r functions_setup}
# relevant barriers taken from theoretical part
initialStockPrice <- 48.68
initialCertificatePrice <- 100
coupon <- 10.15
```

### payoff

The function calculates the payoff at a given valuation date given the stock price.
A direct payoff occurs when price is above the *initialStockPrice*, for each ongoing year the reference price which has to be triggered is lowered by 10% per year relativ to the initial price.

The function adds 10% to start with `index = 1` and to end at 60% of the initial price at maturity.
The values for this equation are taken from the information sheet provided by the bank, as well as the valuation at expiration if none of the previous levels have been triggered.

The calculation for value at expiration is taken from the information sheet.

pricaAtDate = data frame with valuation dates to store results in

```{r payoff_function}
# payoffs at valuation date given the stock prices
payoff <- function(payoff_df) {
  # create vector to store values in
  payoffResults <- vector(mode = "numeric", length = N)
  # loop over all N columns from simulation matrix
  for (i in 1:N) {
    # 1st evaluation date
    if (payoff_df[1, i] >= initialStockPrice) {
      payoffResults[i] <- initialCertificatePrice + coupon * 1
      next
    }
    # 2nd evaluation date
    else if (payoff_df[2, i] >= initialStockPrice * (1 - 0.1)) {
      payoffResults[i] <- initialCertificatePrice + coupon * 2
      next
    }
    # 3rd evaluation date
    else if (payoff_df[3, i] >= initialStockPrice * (1 - 0.2)) {
      payoffResults[i] <- initialCertificatePrice + coupon * 3
      next
    }
    # 4th evaluation date
    else if (payoff_df[4, i] >= initialStockPrice * (1 - 0.3)) {
      payoffResults[i] <- initialCertificatePrice + coupon * 4
      next
    }
    # maturity data
    else if (payoff_df[5,i] >= initialStockPrice * (1 - 0.4)) {
      payoffResults[i] <- initialCertificatePrice + coupon * 5
      next
    }
    # valuation at maturity if no redemption level has been reached
    else {
      payoffResults[i] <-
        payoff_df[5, i] / initialStockPrice * initialCertificatePrice
    }
  }
  return(payoffResults)
}
```

### geometric Brownian Motion

The possible paths of the certificate are simulated using a brownian motion.
Using this method with a sufficient numbers of repetitions, the average value is centred around the price of the certificate. The created paths are put back into the payoff function and are evaluated there.

```{r brown_sim}
N = 1000  # times of repetition for paths

simulation <- function(
    startPrice,  # Stock price at start of Simulation
    riskFreeDaily,  # svensson risk free daily rate
    vola,  # volatility
    T0,  # time remaining until maturity/dates to simulate
    N,  # number of paths
    dT = 1./365
){
    # create matrix of randomly drawn numbers to make vectorizing possible
      # use seed to keep randomization consistent
    set.seed(123)
  randomMatrix <- matrix(rnorm(N * T0), ncol = N, nrow = T0)
    # Brownian Motion into Price Paths
  BM <- exp((riskFreeDaily - vola * vola / 2) * dT + vola * randomMatrix * sqrt(dT))
  BM <- apply(rbind(rep(startPrice, N), BM), 2, cumprod)
  return(BM)
}
```

```{r extra_data}
  # define days until maturity
daysToMaturity <- vector(mode = "numeric", length = length(valuation))
for (i in 1:length(valuation)) {
  daysToMaturity[i] <-
    sum(!weekdays(seq(valuation[i], maturity, "days")) %in% c("Saturday", "Sunday"))
}

# create subset for simulation purposes
dataSim <- data[(if (nrow(data) > traidingDays[1])
  nrow(data) - traidingDays[1]
  else
    1):nrow(data), c("Date", "Stock", "riskFreeDaily", "volaY1", "volaY2", "untilmaturity", "Certificate")]
dataSim$Sim1Y <- "empty"
dataSim$Sim2Y <- "empty"
dataSim$SimImplVola <- "empty"
rownames(dataSim) <- NULL

write_csv(dataSim, "./data/dataSim.csv", col_names = TRUE)

rm(
  maturity,
  emission
)
```

# 1 year volatility

pricing the Certificate using the rolling 1-year volatility of stock returns

```{r 1-year-vola}
for (i in 1:nrow(dataSim)) {
  # define required parameters
  startPrice <- unlist(dataSim[i, "Stock"])
  riskFreeDaily <- unlist(dataSim[i, "riskFreeDaily"])
  vola <- unlist(dataSim[i, "volaY1"])
  T0 <- unlist(dataSim[i, "untilmaturity"])
  # simulate paths using Brownian Motion using previously defined function
  simulatedPaths <- simulation(startPrice, riskFreeDaily, vola, T0, N)
      # reverse matrix for easier indexing
  simulatedInverse <- apply(simulatedPaths, 2, rev)
  
  # store values used as index in matrix
  pathsResult <- matrix(data = NA, nrow = length(daysToMaturity), ncol = N)

  # extract rows for evaluation dates
    # take number to maturity as index
  for (k in 1:length(daysToMaturity)){
    # for every valuation date possible store row in extra matrix
    if (daysToMaturity[k] < nrow(simulatedInverse))
      pathsResult[k,] <- simulatedInverse[daysToMaturity[k],]
    # otherwise put 0 in to fail at payoff function
    else
      pathsResult[k,] <- 0
  }
  # build certificate price as mean discounted payoff from stock prices
  dataSim$Sim1Y[[i]] <-
    as.numeric(exp(-riskFreeDaily * (T0/240)) * mean(payoff(pathsResult), na.rm = TRUE))
}
dataSim$Sim1Y <- as.numeric(dataSim$Sim1Y)
```

# 2 year volatility

Utilise the same function as before, with the only differences being the different `vola` variable, as well as the different column to store the values in.

```{r 2-year-vola}
for (i in 1:nrow(dataSim)) {
  # define required parameters
  startPrice <- unlist(dataSim[i, "Stock"])
  riskFreeDaily <- unlist(dataSim[i, "riskFreeDaily"])
  vola <- unlist(dataSim[i, "volaY2"])
  T0 <- unlist(dataSim[i, "untilmaturity"])
  # simulate paths using Brownian Motion using previously defined function
  simulatedPaths <- simulation(startPrice, riskFreeDaily, vola, T0, N)
      # reverse matrix for easier indexing
  simulatedInverse <- apply(simulatedPaths, 2, rev)
  
  # store values used as index in matrix
  pathsResult <- matrix(data = NA, nrow = length(daysToMaturity), ncol = N)

  # extract rows for evaluation dates
    # take number to maturity as index
  for (k in 1:length(daysToMaturity)){
    # for every valuation date possible store row in extra matrix
    if (daysToMaturity[k] < nrow(simulatedInverse))
      pathsResult[k,] <- simulatedInverse[daysToMaturity[k],]
    # otherwise put 0 in to fail at payoff function
    else
      pathsResult[k,] <- 0
  }
  # build certificate price as mean discounted payoff from stock prices
  dataSim$Sim2Y[[i]] <-
    as.numeric(exp(-riskFreeDaily * (T0/240)) * mean(payoff(pathsResult), na.rm = TRUE))
}
dataSim$Sim2Y <- as.numeric(dataSim$Sim2Y)
```

remove unnecessary variables

```{r rm_1}
rm(
  startPrice,
  riskFreeDaily,
  T0,
  vola,
  pathsResult,
  simulatedInverse,
  simulatedPaths
)
# manually collect garbage
gc()
```

# Visualization

### simulated vs real prices

```{r sim_vs_real_plot}
ggplot(
  data = dataSim,
  aes(x = Date)) +
  geom_line(aes(y = Certificate, color = 'Market Value')) +
  geom_line(aes(y = as.numeric(Sim1Y), color = 'Simulation - 1 year volatility')) +
  geom_line(aes(y = as.numeric(Sim2Y), color = 'Simulation - 2 year volatility')) +
  theme_bw(base_size = 25) +
  theme(legend.position = c(0.8, 0.8),
        legend.title = element_blank()) +
  ggtitle("Valuation") +
  xlab("Time") +
  ylab("Price in EUR")

ggsave(filename = "./figures/CertificateValuation.png", dpi = 320, width = 21, height = 7)
```

# Error statistics
 
 Define all required error terms. These are root mean squared error (RMSE), mean absolute error (MAE) and error quantiles.

```{r error_terms}
  # root mean squared error
RMSE <- function(predictedValue, trueValue){
  sqrt(mean((predictedValue - trueValue)^2))
}
  # mean absolute error
MAE <- function(predictedValue, trueValue){
  mean(abs(predictedValue - trueValue))
}
  # quantiles
ErrorQuantile <- function(predictedValue, trueValue){
  quantile(abs(predictedValue - trueValue), probs = c(.25, .5, .75))
}
```
 
### RMSE
 
```{r rmse}
# 1 year volatility
RMSE(dataSim$Sim1Y, dataSim$Certificate)

# 2 year volatility
RMSE(dataSim$Sim2Y, dataSim$Certificate)
```

### MAE
 
```{r mae}
# 1 year volatility
MAE(dataSim$Sim1Y, dataSim$Certificate)

# 2 year volatility
MAE(dataSim$Sim2Y, dataSim$Certificate)
```

### Error Quantiles
 
```{r quantiles}
# 1 year volatility
ErrorQuantile(dataSim$Sim1Y, dataSim$Certificate)

# 2 year volatility
ErrorQuantile(dataSim$Sim2Y, dataSim$Certificate)
```
