---
title: "performance Analysis"
author: "Jan Besler"
date: "`r Sys.Date()`"
output: html_document
---

# Setup

> ***libraries***

before starting the work space is cleared and all necessary libraries are loaded.

```{r setup, include=FALSE}
# load libraries
library(knitr)      # use .Rmd files

  # reading data
library(readr)      # reading csv

  # manipulate data2
library(tidyr)      # manipulating data2
library(dplyr)      # manipulating data2
library(rlist)      # working with lists
library(stringr)    # work with strings
library(tibble)     # create data2 frames
library(stringi)    # process charactar strings
library(lubridate)  # transform character into dates

  # statistics
library(Matrix)     # required for lme4
library(moments)    # skewness & curtosis

  # visuals
library(ggplot2)    # make graphs
library(ggExtra)    # additional functionality for ggplot2

knitr::opts_chunk$set(echo = TRUE)
```

> For this task, assume a fictive investor with an investment horizon of T* = 1 year, planning to invest 10,000 euros. Assume that the investor has the possibility to invest in a stock or index, and put options for portfolio insurance purposes.

# Part VI

**Performance Analysis: without Risk Management**

Choose an appropriate simulation approach to derive the payoff/return distribution at
the end of the investment horizon T* of an investment in the stock/index. Calculate
meaningful performance and risk measures and explain.

The chosen simulation approach is a Geometric Brownian Motion, since this has been already been used for the previous parts of the assignment.

### functions

```{r VI_0}
simulation <- function(
    startPrice,  # Stock price at start of Simulation
    riskFreeDaily,  # svensson risk free daily rate
    vola,  # volatility
    T0,  # time remaining until maturity/dates to simulate
    N,  # number of paths
    dT = 1./365
){
    # create matrix of randomly drawn numbers to make vectorizing possible
      # use seed to keep randomization consistent
    set.seed(123)
  randomMatrix <- matrix(rnorm(N * T0), ncol = N, nrow = T0)
    # Brownian Motion into Price Paths
  BM <- exp((riskFreeDaily - vola * vola / 2) * dT + vola * randomMatrix * sqrt(dT))
  BM <- apply(rbind(rep(startPrice, N), BM), 2, cumprod)
  return(BM)
}

traidingDays <- c(240, 480)
```

### data & data frames

At first the payoff distribution is calculated, using the already defined functions and data2 sets.

```{r VI_1}
# load Stock data2 of Morphosys AG & svensson parameters
data2 <- read_csv("./data/dataPart2.csv",
                 col_select = c(
                   "Date",
                   "Stock",
                   "volaY1",
                   "volaY2",
                   "riskFreeDaily"
                 ))
# only keep newest date as start for simulation
data2 <- data2[nrow(data2),]
# set number of paths for brownian motion
N = 1000

# create table to store all results in.
VariableInvest <- as.data.frame(matrix(data = NA,
                     ncol = 6,  # position, mean return Underlying, vola, skewness, VaR-95, VaR-99
                     nrow = 5))  # % of portfolio in position
  # rename columns
VariableInvest <- VariableInvest %>% rename(
  'investment in Puts' = V1,
  'mean return' = V2,
  volatility = V3,
  skewness = V4,
  'VaR 95 %' = V5,
  'VaR 99 %' = V6
)
  # fill position column
VariableInvest$`investment in Puts` <- c("0 %", "3 %", "6 %", "9 %", "12 %")

# create table to store all results in.
VariableStrike <- as.data.frame(matrix(data = NA,
                     ncol = 6,  # position, mean return Underlying, vola, skewness, VaR-95, VaR-99
                     nrow = 10))  # number of Strike Prices + none
  # rename columns
VariableStrike <- VariableStrike %>% rename(
  'Strike Price' = V1,
  'mean return' = V2,
  volatility = V3,
  skewness = V4,
  'VaR 95 %' = V5,
  'VaR 99 %' = V6
)
  # fill position column
VariableStrike[1:nrow(VariableStrike),"Strike Price"] <- seq(17.5, 22, .5)


# amount of stocks and puts for different scenarios
InvestmentScenarios <- as.data.frame(matrix(data = NA,
                                            ncol = 5,
                                            nrow = nrow(VariableInvest)))
InvestmentScenarios <- InvestmentScenarios %>% rename(
  "Stocks in Percent" = V1,
  "Puts in Percent" = V2,
  AmountStocks = V3,
  AmountPuts = V4,
  Cash = V5
)
InvestmentScenarios$`Stocks in Percent` <- c(100, 97, 94, 91, 88)
InvestmentScenarios$`Puts in Percent` <- rep(100, 5) - InvestmentScenarios$`Stocks in Percent`

# matrix to store scenarios in
PortfolioReturnMatrix <- as.data.frame(matrix(data = NA, ncol = 3, nrow = N))
PortfolioReturnMatrix <- PortfolioReturnMatrix %>% rename(
  withoutPuts = V1,
  StrikePrice = V2,
  PutInvestment = V3
)
```

### Return & Payoff

```{r VI_2}
# define required parameters
startPrice <- unlist(data2[nrow(data2),"Stock"])
riskFreeDaily <- unlist(data2[nrow(data2),"riskFreeDaily"])
vola <- unlist(data2[nrow(data2),"volaY1"])
T0 <- traidingDays[1]

# simulate paths using Brownian Motion using previously defined function
simulatedPaths <- simulation(startPrice, riskFreeDaily, vola, T0, N)


# absolut difference/ payoff
StockPayoffVector <- vector(mode = "numeric", length = N)
StockPayoffVector <- (simulatedPaths[nrow(simulatedPaths),] - rep(startPrice, N)) 
  # turn into data2 frame with one column & rename column
StockPayoffMatrix <- as.data.frame(StockPayoffVector)
StockPayoffMatrix <- rename(StockPayoffMatrix,
                            payoff = StockPayoffVector)

# Return
StockReturnVector <- vector(mode = "numeric", length = N)
StockReturnVector <- (simulatedPaths[nrow(simulatedPaths),] - rep(startPrice, N)) / rep(startPrice, N)
  # turn into data2 frame with one column & rename column
StockReturnMatrix <- as.data.frame(StockReturnVector)
StockReturnMatrix <- rename(StockReturnMatrix,
                            Return = StockReturnVector)

PortfolioReturnMatrix$withoutPuts <- StockReturnVector
```



### performance and risk measures

At last meaningful measures regarding the performance and risk measures are calculated. These include the mean return for performance and for risk measures the volatility, skewness and Value at Risk for 95% and 99%.

The Value-at-Risk (VaR) provides an aggregate measure of risk by assessing worst-case scenarios. The historical (non-parametric) VaR is used.

##### functions

```{r VI_4}
  # annual volatility - input is vector
volatilityFunction <- function(PortfolioReturn) {
  vola <- sqrt(var(PortfolioReturn)) * 100  # in percent
  return(vola)
}

# value at risk - input is vector
VaRFunction <- function(PortfolioReturn) {
  # matrix to store all VaR in
  VaRResult <- vector(mode = "numeric", length = 2)  # 2 elements for 95 & 99 VaR
  # order input vector
  VaRReturns <- sort(PortfolioReturn, decreasing = TRUE)
  # pick quantiles percentile
  VaRResult <- quantile(VaRReturns, probs = c(.05, .01)) * rep(100, 2)  # in percent
  return(VaRResult)
}
```

### Visualization

In the second step the data2 is visualised as a Kernel Density Plot with a Histogram underneath to better show where the data2 is located.

```{r VI_2, warning=FALSE}
ggplot(
  data = StockPayoffMatrix,
  aes(x = payoff)) +
  geom_histogram(aes(y = ..density..),
                 bins = 100,
                 color = 'darkblue',
                 fill = 'white') +
  geom_density(color = 'red',
               fill = 'red',
               alpha = 0.3,
               linetype = 'dashed') +
  labs(x = "Payoff in EUR",
      y = "density") +
  ggtitle("Density estimation of Portfolio Payoff") +
  theme_minimal()

# save the graph
ggsave(
  "./figures/PayoffDistribution.png",
  dpi = 320
)

ggplot(data = StockReturnMatrix,
       aes(x = Return)) +
  geom_histogram(aes(y = ..density..),
                 bins = 100,
                 color = 'darkblue',
                 fill = 'white') +
  geom_density(
    color = 'red',
    fill = 'red',
    alpha = 0.3,
    linetype = 'dashed'
  ) +
  labs(x = "Return in %",
       y = "density") +
  ggtitle("Density estimation of Portfolio Return") +
  theme_bw(base_size = 25)

# save the graph
ggsave(
  "./figures/ReturnDistribution.png",
  dpi = 320,
  width = 21,
  height = 10
)
```


# Part VII

**Performance Analysis: with Risk Management**

Implement a portfolio insurance strategy by including put options to hedge the portfolio against downside risks. Consider fictitious puts, which prices are based on the historical volatility of the underlying, and analyze the impact of

  - the fraction of initial wealth invested in put options, and
  - the strike K of the put options

on the performance and risk measures of the resulting portfolio. Illustrate and discuss your observations and compare them to the results of Problem VI.

### functions

```{r VII_1}
# define Black Scholes Function for Put Pricing
BlackScholes <-
  function(StockPrice,
           StrikePrice,
           RiskFreeRate,
           TimeToExpiration,
           volatility) {
    d1 <-
      (log(StockPrice / StrikePrice) + (RiskFreeRate + volatility ^ 2 / 2) *
         TimeToExpiration) / (volatility * sqrt(TimeToExpiration))
    d2 <- d1 - volatility * sqrt(TimeToExpiration)
    value <-
      (
        StrikePrice * exp(-RiskFreeRate * TimeToExpiration) * pnorm(-d2) - StockPrice *
          pnorm(-d1)
      )
    return(value)
  }

# claculate volatility as log return - input single price (not first one though)
logReturn <- function(StockPrice){
  logVola <- sd(log(StockPrice[[-1]]) - log(StockPrice))  # daily stock return
}
```

### calculations

The time variant values such as volatility stay in their daily format, as the TimeToExpiration is also given on a daily basis.
The volatility is calculated as the standard deviation of the returns for the given time frame. Here the simulated returns from Part VI are taken again.
Since there is no riskfree Rate for the simulated paths, the average from the data used in for the Certificate is used as a mean as an approximation.

```{r VII_2}
# build fictitious put option
StrikePriceSample <- as.numeric(VariableStrike[1:nrow(VariableStrike), "Strike Price"])
  # create matrix to store results
PutMatrix <- matrix(data = NA,
                     ncol = 4,
                     nrow = length(StrikePriceSample))
PutMatrix[,1] <- StrikePriceSample
colnames(PutMatrix) <- c("StrikePrice", "PriceBegininng", "PriceMaturity", "Return")

  # calculate Put Option Prices at beginning of time window
for(i in 1:length(StrikePriceSample)){
  PutMatrix[i,2] <- BlackScholes(
    StockPrice = unlist(data2[1, "Stock"]),
    StrikePrice = unlist(StrikePriceSample[i]),
    RiskFreeRate = unlist(data2[1, "riskFreeDaily"]),
    TimeToExpiration = traidingDays[1],
    volatility = unlist(data2[nrow(data2), "volaY1"]) / sqrt(traidingDays[1]))[1]
}
  # calculate Put Option Prices at maturity
    # create matrix to store values in
PutMaturity <- matrix(data = NA, ncol = N, nrow = length(StrikePriceSample))
    # riskfree Rate is approximated as the mean of all available dates from Part 1
dataSim <- read_csv("./data/dataSim.csv")

    # volatility as log return from Stock Prices
vola1 <- vector(mode = "numeric", length = nrow(simulatedPaths)-1)
diff <- vector(mode = "numeric", length = nrow(simulatedPaths)-1)
volaSimulated <- matrix(data = NA, ncol = N, nrow = nrow(simulatedPaths)-1)
# calculate the volatility for 1 year
for(j in 1:N) {
  for (i in 2:nrow(simulatedPaths)) {
    diff[i - 1] <-
      log(simulatedPaths[i - 1, j]) - log(simulatedPaths[i, j])  # daily stock return
    vola1[i - 1] <-
      sd(diff) * sqrt(traidingDays[1])  # standard deviation of stock return for year 1
  }
  volaSimulated[,j] <- vola1
}

for(i in 1:length(StrikePriceSample)){
  for(j in 1:N){
    PutMaturity[i,j] <- BlackScholes(
      StockPrice = unlist(simulatedPaths[nrow(simulatedPaths), j]),
      StrikePrice = unlist(StrikePriceSample[i]),
      RiskFreeRate = mean(dataSim$riskFreeDaily, na.rm = TRUE),
      TimeToExpiration = .001,  # approximation to zero
      volatility = unlist(volaSimulated[nrow(volaSimulated),j])
      )[1]
  }
}

  # calculate Gain and Loss for Put Options
for(i in 1:length(StrikePriceSample)){
    # calculate mean price of Put at Maturity 
  PutMatrix[i,3] <- mean(PutMaturity[,i])
  PutMatrix[i,4] <- PutMatrix[i,3] - PutMatrix[i,2]
}

# fill InvestmentScenario data frame
for(i in 1:nrow(InvestmentScenarios)) {
  InvestmentScenarios[i,"AmountStocks"] <-
    floor(10000 * InvestmentScenarios[i, "Stocks in Percent"]/100 / data2$Stock)
  InvestmentScenarios[i,"AmountPuts"] <-
    floor(10000 * InvestmentScenarios[i, "Puts in Percent"]/100 / PutMatrix[PutMatrix[,"StrikePrice"] == 19,"PriceBegininng"])
  InvestmentScenarios[i,"Cash"] <- 10000 - (InvestmentScenarios[i,"AmountStocks"] * data2$Stock + InvestmentScenarios[i,"AmountPuts"] * PutMatrix[PutMatrix[,"StrikePrice"] == 19,"PriceBegininng"])
}
```


### Portfolio

##### variable Puts

varying the percentage of Puts in the investment & holding Strike Price constant.
Strike price is chosen to be 19 (StrikePriceSample[4])

```{r VII_3}
# calculate portfolio value (positions * price) at maturity - different for each N
PortfolioMaturity <- matrix(data = NA, ncol = N, nrow = nrow(InvestmentScenarios))
for(j in 1:N){
  for (i in 1:nrow(InvestmentScenarios)) {
    PortfolioMaturity[i,j] <-
      simulatedPaths[nrow(simulatedPaths),j] * InvestmentScenarios[i, "AmountStocks"] +
      PutMaturity[4,j] * InvestmentScenarios[i, "AmountPuts"] +
      InvestmentScenarios[i, "Cash"]
  }
}

# mean return from both portfolio values
PortfolioReturn = matrix(data = NA, ncol = N, nrow = nrow(InvestmentScenarios))
for(i in 1:nrow(PortfolioReturn)){
  PortfolioReturn[i,] <- (PortfolioMaturity[i,] - rep(10000, N)) / rep(10000, N)
}
PortfolioReturnMatrix$PutInvestment <- PortfolioReturn[5,]

# risk measures
for(i in 1:nrow(VariableInvest)){
  # mean return
VariableInvest[i, "mean return"] <- mean(PortfolioReturn[i,])
  # volatility
VariableInvest[i, "volatility"] <- volatilityFunction(PortfolioReturn[i,])
  # skewness
VariableInvest[i, "skewness"] <- skewness(PortfolioReturn[i,])
  # VaR 95%
VariableInvest[i, "VaR 95 %"] <- VaRFunction(PortfolioReturn[i,])[1]
  # VaR 99%
VariableInvest[i, "VaR 99 %"] <- VaRFunction(PortfolioReturn[i,])[2]
}
```

##### variable Strike Price

Strike Price is varied, but Investment in Portfolio is hold constant for 6% of Puts

```{r}
# calculate portfolio value (positions * price) at maturity - different for each N
PortfolioMaturity <- matrix(data = NA, ncol = N, nrow = length(StrikePriceSample))
for(j in 1:N){
  for (i in 1:length(StrikePriceSample)) {
    PortfolioMaturity[i,j] <-
      simulatedPaths[nrow(simulatedPaths),j] * InvestmentScenarios[3, "AmountStocks"] +
      PutMaturity[i,j] * InvestmentScenarios[3, "AmountPuts"] +
      InvestmentScenarios[3, "Cash"]
  }
}

# mean return from both portfolio values
PortfolioReturn = matrix(data = NA, ncol = N, nrow = length(StrikePriceSample))
for(i in 1:nrow(PortfolioReturn)){
  PortfolioReturn[i,] <- (PortfolioMaturity[i,] - rep(10000, N)) / rep(10000, N)
}
PortfolioReturnMatrix$StrikePrice <- PortfolioReturn[2,]

# risk measures
for(i in 1:nrow(VariableStrike)){
  # mean return
VariableStrike[i, "mean return"] <- mean(PortfolioReturn[i,])
  # volatility
VariableStrike[i, "volatility"] <- volatilityFunction(PortfolioReturn[i,])
  # skewness
VariableStrike[i, "skewness"] <- skewness(PortfolioReturn[i,])
  # VaR 95%
VariableStrike[i, "VaR 95 %"] <- VaRFunction(PortfolioReturn[i,])[1]
  # VaR 99%
VariableStrike[i, "VaR 99 %"] <- VaRFunction(PortfolioReturn[i,])[2]
}
```

### Visualization

```{r plot_VII, eval=TRUE}
ggplot(
  data = PortfolioReturnMatrix) +
  geom_density(aes(x = withoutPuts,
               color = 'without Puts')) +
  geom_density(aes(x = StrikePrice,
               color = 'with 18 EUR Strike Price')) +
  geom_density(aes(x = PutInvestment,
               color = 'with 12% Puts')) +
  labs(x = "Return in %",
      y = "density") +
  ggtitle("Density estimation of Portfolio Return",
          "for different scenarios") +
  xlim(-1,1.25) +
  theme_bw(base_size = 25) +
  theme(legend.position = c(0.8, 0.8),
        legend.title = element_blank())

# save the graph
ggsave(
  "./figures/PortfolioScenarios.png",
  dpi = 320,
  width = 21,
  height = 7
)
```
