---
title: "selection characteristics"
author: "Jan Besler"
date: "`r Sys.Date()`"
output: html_document
---

### Setup

> ***libraries***

before starting the work space is cleared and all necessary libraries are loaded.

```{r setup, include=FALSE}
# load libraries
library(knitr)      # use .Rmd files

  # reading data
library(readr)      # reading csv
library(readxl)     # reading excel
library(foreign)    # read in Stata, SPSS & SAS files
library(haven)      # read in Stata, SPSS & SAS files
library(DBI)        # connecting database
library(RSQLite)    # sql support
library(jsonlite)   # using APIs
library(httr)       # using APIs
library(rvest)      # webscraping
library(xml2)       # parse xml code
library(here)       # help with relative paths in r projects

  # manipulate data
library(tidyr)      # manipulating data
library(dplyr)      # manipulating data
library(rlist)      # working with lists
library(stringr)    # work with strings
library(tibble)     # create data frames
library(stringi)    # process charactar strings

  # statistics
library(Matrix)     # required for lme4
library(lme4)       # used for multilevel modeling
library(lfe)        # fixed effect regression

  # visuals
library(ggplot2)    # make graphs
library(ggExtra)    # additional functionality for ggplot2
library(ggpubr)     # combine multiple graphs into one
library(ggrepel)    # describe lines in graph

knitr::opts_chunk$set(echo = TRUE)
```

## Company

> the company of interest is ***MorphoSys AG*** (ISIN:DE0006632003) with the [Express Certificate](https://keyinvest-de.ubs.com/produkt/detail/index/isin/DE000UBS1KS0) (ISIN:DE000UBS1KS0) from UBS

the express certificate has the following early redemption observation dates:

-   24^th^ August 2022 = j1
-   24^th^ August 2023 = j2
-   26^th^ August 2024 = j3
-   25^th^ August 2025 = j4

and the maturity date of 31^st^ August 2026

parameters:

-   kick in level at 60% of emission price
-   emission price 100 €
-   coupon of 10,15 € times year of maturity
-   early redemption level:
    -   j1 = 100 %
    -   j2 = 90 %
    -   j3 = 80 %
    -   j4 = 70 % in reference to intial price

At first the payoff function has to be defined with the parameters taken from a [term sheet](https://keyinvest-de.ubs.com/filedb/deliver/xuuid/g001aa76ff7bc23a9fe6b2eecba79034fc0b/name/DE000UBS1KS0_Endg%C3%BCltige+Bedingungen_en.pdf) from UBS

```{r part_I}
# the values for these parameters are
initial_price <- 48.68
barrier <- 0.6*initial_price
coupon <- 10.15

# define the theoretical payoff function
payoff <- function(p, year) {
  if (p < (initial_price * (1.1 - year*0.1)))
    p
  else #(p >= initial_price * (1.1 - year*0.1))
    100 + year*coupon
}

MORExZ <- data.frame(stock_price = c(1:175),
                     payoff_y1 = NA,
                     payoff_y2 = NA,
                     payoff_y3 = NA,
                     payoff_y4 = NA,
                     payoff_maturity = NA)
for(j in 1:nrow(MORExZ)) {
  for (i in 1:5) {
    MORExZ[j, i + 1] <- payoff(p = MORExZ$stock_price[[j]], year = i)
  }
}
# change x values to be closer to previous values at observation dates
MORExZ[30,1] <- 29.001
MORExZ[35,1] <- 34.001
MORExZ[39,1] <- 38.001
MORExZ[44,1] <- 43.001
MORExZ[49,1] <- 48.001
```

```{r part_II}
# visualization of payoff
ggplot(data = MORExZ,
       aes(x = stock_price)) +
  # stock price line
  geom_line(aes(y = stock_price,
                color = 'underlying'),
            linetype = 'twodash') +
  # payoff at j1 (first early payoff)
  geom_line(aes(y = payoff_y1,
                color = 'early payoff year 1'),
            linetype = 'longdash') +
  # payoff at j2 (second early payoff)
  geom_line(aes(y = payoff_y2,
                color = 'early payoff year 2'),
            linetype = 'longdash') +
  # payoff at j3 (third early payoff)
  geom_line(aes(y = payoff_y3,
                color = 'early payoff year 3'),
            linetype = 'longdash') +
  # payoff at j4 (third early payoff)
  geom_line(aes(y = payoff_y4,
                color = 'early payoff year 4'),
            linetype = 'longdash') +
  # derivitive payoff at maturity
  geom_line(aes(y = payoff_maturity,
                color = 'payoff maturity'),
            linetype = 'solid') +
  # naming the axis and legend
  ylab("derivative payoff (EUR)") +
  xlab("stock price (EUR)") +
  theme_bw(base_size = 25) +
  theme(legend.position = c(0.85, 0.25),
        legend.title = element_blank()) +
  ggtitle("Payoff structure of Express Certificate")

# save the figure
ggsave(filename = "./figures/theoryPayoff.png", dpi = 320, width = 21, height = 7)
```

```{r part_I_II}
rm(
  theoryPayoff,
  MORExZ
)
```