---
title: "performance Analysis"
author: "Jan Besler"
date: "`r Sys.Date()`"
output: html_document
---

# Setup

> ***libraries***

before starting the work space is cleared and all necessary libraries are loaded.

```{r setup, include=FALSE}
# load libraries
library(knitr)      # use .Rmd files

  # reading data
library(readr)      # reading csv

  # manipulate data
library(tidyr)      # manipulating data
library(dplyr)      # manipulating data
library(rlist)      # working with lists
library(stringr)    # work with strings
library(tibble)     # create data frames
library(stringi)    # process charactar strings
library(lubridate)  # transform character into dates

  # statistics
library(Matrix)     # required for lme4

  # visuals
library(ggplot2)    # make graphs
library(ggExtra)    # additional functionality for ggplot2

knitr::opts_chunk$set(echo = TRUE)

# replicate data frames
PortfolioReturnMatrixModified <- PortfolioReturnMatrix
PutMatrixModified <- PutMatrix
VariableInvestModified <- VariableInvest
VariableStrikeModified <- VariableStrike

# data frame to hold all changed variables
FinalTable <- as.data.frame(matrix(data = NA,
                     ncol = 7,  # properties, case, mean return, vola, skewness, VaR-95, VaR-99
                     nrow = 12))  # 4*3 rows
  # rename columns
FinalTable <- FinalTable %>% rename(
  Properties = V1,
  Case = V2,
  'mean return' = V3,
  volatility = V4,
  skewness = V5,
  'VaR 95%' = V6,
  'VaR 99%' = V7
)
  # fill position column
FinalTable$Properties <- rep(c("without hedging", "12% Puts", "Strike Price = 18"),4)
FinalTable$Case <- c(rep("base case",3),rep("-5% volatility",3), rep("-20% stock price",3),rep("+5% volatility & -20% stock price",3))

FinalTable[1,3:7] <- VariableInvest[1,2:6]
FinalTable[2,3:7] <- VariableInvest[5,2:6]
FinalTable[3,3:7] <- VariableStrike[2,2:6]
```

# Part VIII

For a specific portfolio allocation from Problem VII, consider that the volatility for pricing
purposes of the options equals the historical volatility of the underlying ±5 percentage
points and/or a sudden drop of the stock/index price by 20% after half a year. Compare
the results and explain.

Exemplary only one case from variable Investment and variable Strike Price are compared plus the base case.

### Case 1

> A 5 % drop in volatility. 

##### Simulation

At first the simulation has to be modified. For the riskfree Rate the same value is approximated as before in Task VII.
The simulated values are cut in half and resumed from the cut part.

```{r VIII_1}
originalSimulation <- simulation(startPrice, riskFreeDaily, vola, T0, N)[1:floor(nrow(simulatedPaths)/2),]

# create for each N one single simulation and add it back
ResultMatrix <- matrix(data = NA, ncol = N, nrow = (nrow(simulatedPaths) - nrow(originalSimulation) + 1))
for(j in 1:N){
  ResultMatrix[,j] <- simulation(
    startPrice = unlist(originalSimulation[nrow(originalSimulation),j]),
    riskFreeDaily = mean(dataSim$riskFreeDaily, na.rm = TRUE),
    vola = unlist(volaSimulated[nrow(originalSimulation),j] - 0.05),  # minus 5% vola
    T0 = nrow(simulatedPaths) - nrow(originalSimulation),
    N = 1
  )
}
modifiedSimulation <- rbind(originalSimulation, ResultMatrix[2:nrow(ResultMatrix),])
```

##### Puts

only prices for maturity have to be calculated newly.

```{r VIII_2}
    # create matrix to store values in
PutMaturityModified <- matrix(data = NA, ncol = N, nrow = length(StrikePriceSample))
    # riskfree Rate is approximated as the mean of all available dates from Part 1
dataSim <- read_csv("./data/dataSim.csv")

    # volatility as log return from Stock Prices
vola1 <- vector(mode = "numeric", length = nrow(modifiedSimulation)-1)
diff <- vector(mode = "numeric", length = nrow(modifiedSimulation)-1)
volaSimulated <- matrix(data = NA, ncol = N, nrow = nrow(modifiedSimulation)-1)
# calculate the volatility for 1 year
for(j in 1:N) {
  for (i in 2:nrow(modifiedSimulation)) {
    diff[i - 1] <-
      log(modifiedSimulation[i - 1, j]) - log(modifiedSimulation[i, j])  # daily stock return
    vola1[i - 1] <-
      sd(diff) * sqrt(traidingDays[1])  # standard deviation of stock return for year 1
  }
  volaSimulated[,j] <- vola1
}

for(i in 1:length(StrikePriceSample)){
  for(j in 1:N){
    PutMaturityModified[i,j] <- BlackScholes(
      StockPrice = unlist(modifiedSimulation[nrow(modifiedSimulation), j]),
      StrikePrice = unlist(StrikePriceSample[i]),
      RiskFreeRate = mean(dataSim$riskFreeDaily, na.rm = TRUE),
      TimeToExpiration = .001,  # approximation to zero
      volatility = unlist(volaSimulated[nrow(volaSimulated),j])
      )[1]
  }
}
  # calculate Gain and Loss for Put Options
for(i in 1:length(StrikePriceSample)){
    # calculate mean price of Put at Maturity 
  PutMatrixModified[i,3] <- mean(PutMaturityModified[,i])
  PutMatrixModified[i,4] <- PutMatrixModified[i,3] - PutMatrixModified[i,2]
}
```


### Portfolio

##### variable Strike Price

```{r VIII_3}
# calculate portfolio value (positions * price) at maturity - different for each N
PortfolioMaturityModified <- matrix(data = NA, ncol = N, nrow = nrow(InvestmentScenarios))
for(j in 1:N){
  for (i in 1:nrow(InvestmentScenarios)) {
    PortfolioMaturityModified[i,j] <-
      modifiedSimulation[nrow(modifiedSimulation),j] * InvestmentScenarios[i, "AmountStocks"] +
      PutMaturityModified[4,j] * InvestmentScenarios[i, "AmountPuts"] +
      InvestmentScenarios[i, "Cash"]
  }
}

# mean return from both portfolio values
PortfolioReturnModified = matrix(data = NA, ncol = N, nrow = nrow(InvestmentScenarios))
for(i in 1:nrow(PortfolioReturnModified)){
  PortfolioReturnModified[i,] <- (PortfolioMaturityModified[i,] - rep(10000, N)) / rep(10000, N)
}
PortfolioReturnMatrixModified$PutInvestment <- PortfolioReturnModified[5,]

# risk measures
for(i in 1:nrow(VariableInvestModified)){
  # mean return
VariableInvestModified[i, "mean return"] <- mean(PortfolioReturnModified[i,])
  # volatility
VariableInvestModified[i, "volatility"] <- volatilityFunction(PortfolioReturnModified[i,])
  # skewness
VariableInvestModified[i, "skewness"] <- skewness(PortfolioReturnModified[i,])
  # VaR 95%
VariableInvestModified[i, "VaR 95 %"] <- VaRFunction(PortfolioReturnModified[i,])[1]
  # VaR 99%
VariableInvestModified[i, "VaR 99 %"] <- VaRFunction(PortfolioReturnModified[i,])[2]
}

FinalTable[4,3:7] <- VariableInvestModified[1,2:6]
FinalTable[5,3:7] <- VariableInvestModified[5,2:6]
```

##### variable Strike Price

Strike Price is varied, but Investment in Portfolio is hold constant for 6% of Puts

```{r}
# calculate portfolio value (positions * price) at maturity - different for each N
PortfolioMaturityModified <- matrix(data = NA, ncol = N, nrow = length(StrikePriceSample))
for(j in 1:N){
  for (i in 1:length(StrikePriceSample)) {
    PortfolioMaturityModified[i,j] <-
      modifiedSimulation[nrow(modifiedSimulation),j] * InvestmentScenarios[3, "AmountStocks"] +
      PutMaturityModified[i,j] * InvestmentScenarios[3, "AmountPuts"] +
      InvestmentScenarios[3, "Cash"]
  }
}

# mean return from both portfolio values
PortfolioReturnModified = matrix(data = NA, ncol = N, nrow = length(StrikePriceSample))
for(i in 1:nrow(PortfolioReturnModified)){
  PortfolioReturnModified[i,] <- (PortfolioMaturityModified[i,] - rep(10000, N)) / PortfolioMaturityModified[i,]
}
PortfolioReturnMatrixModified$StrikePrice <- PortfolioReturnModified[2,]

# risk measures
for(i in 1:nrow(VariableStrikeModified)){
  # mean return
VariableStrikeModified[i, "mean return"] <- mean(PortfolioReturnModified[i,])
  # volatility
VariableStrikeModified[i, "volatility"] <- volatilityFunction(PortfolioReturnModified[i,])
  # skewness
VariableStrikeModified[i, "skewness"] <- skewness(PortfolioReturnModified[i,])
  # VaR 95%
VariableStrikeModified[i, "VaR 95 %"] <- VaRFunction(PortfolioReturnModified[i,])[1]
  # VaR 99%
VariableStrikeModified[i, "VaR 99 %"] <- VaRFunction(PortfolioReturnModified[i,])[2]
}

FinalTable[6,3:7] <- VariableStrikeModified[2,2:6]
```

### Visualization

```{r plot_VIII_case1, eval=TRUE}
ggplot(
  data = PortfolioReturnMatrixModified) +
  geom_density(aes(x = withoutPuts,
               color = 'without Puts')) +
  geom_density(aes(x = StrikePrice,
               color = 'with 18 EUR Strike Price')) +
  geom_density(aes(x = PutInvestment,
               color = 'with 12% Puts')) +
  labs(x = "Return in EUR",
      y = "density") +
  ggtitle("Density estimation of Portfolio Return",
          "-5% volatility") +
  xlim(-1,1) +
  theme_bw(base_size = 25) +
  theme(legend.position = c(0.8, 0.8),
        legend.title = element_blank())

# save the graph
ggsave(
  "./figures/PartVIII_case1.png",
  dpi = 320,
  width = 11,
  height = 7
)
```

### Case 2

> A 20 % drop in the stock price. 

##### Simulation

At first the simulation has to be modified. For the riskfree Rate the same value is approximated as before in Task VII.
The simulated values are cut in half and resumed from the cut part.

```{r VIII_1}
# create for each N one single simulation and add it back
for(j in 1:N){
  ResultMatrix[,j] <- simulation(
    startPrice = unlist(originalSimulation[nrow(originalSimulation),j]  * 0.8),  # 20% percent drop
    riskFreeDaily = mean(dataSim$riskFreeDaily, na.rm = TRUE),
    vola = unlist(volaSimulated[nrow(originalSimulation),j]),
    T0 = nrow(simulatedPaths) - nrow(originalSimulation),
    N = 1
  )
}
modifiedSimulation <- rbind(originalSimulation, ResultMatrix[2:nrow(ResultMatrix),])
```

##### Puts

only prices for maturity have to be calculated newly.

```{r VIII_2}
    # create matrix to store values in
PutMaturityModified <- matrix(data = NA, ncol = N, nrow = length(StrikePriceSample))
    # riskfree Rate is approximated as the mean of all available dates from Part 1
dataSim <- read_csv("./data/dataSim.csv")

    # volatility as log return from Stock Prices
vola1 <- vector(mode = "numeric", length = nrow(modifiedSimulation)-1)
diff <- vector(mode = "numeric", length = nrow(modifiedSimulation)-1)
volaSimulated <- matrix(data = NA, ncol = N, nrow = nrow(modifiedSimulation)-1)
# calculate the volatility for 1 year
for(j in 1:N) {
  for (i in 2:nrow(modifiedSimulation)) {
    diff[i - 1] <-
      log(modifiedSimulation[i - 1, j]) - log(modifiedSimulation[i, j])  # daily stock return
    vola1[i - 1] <-
      sd(diff) * sqrt(traidingDays[1])  # standard deviation of stock return for year 1
  }
  volaSimulated[,j] <- vola1
}

for(i in 1:length(StrikePriceSample)){
  for(j in 1:N){
    PutMaturityModified[i,j] <- BlackScholes(
      StockPrice = unlist(modifiedSimulation[nrow(modifiedSimulation), j]),
      StrikePrice = unlist(StrikePriceSample[i]),
      RiskFreeRate = mean(dataSim$riskFreeDaily, na.rm = TRUE),
      TimeToExpiration = .001,  # approximation to zero
      volatility = unlist(volaSimulated[nrow(volaSimulated),j])
      )[1]
  }
}
  # calculate Gain and Loss for Put Options
for(i in 1:length(StrikePriceSample)){
    # calculate mean price of Put at Maturity 
  PutMatrixModified[i,3] <- mean(PutMaturityModified[,i])
  PutMatrixModified[i,4] <- PutMatrixModified[i,3] - PutMatrixModified[i,2]
}
```


### Portfolio

##### variable Strike Price

```{r VIII_3}
# calculate portfolio value (positions * price) at maturity - different for each N
PortfolioMaturityModified <- matrix(data = NA, ncol = N, nrow = nrow(InvestmentScenarios))
for(j in 1:N){
  for (i in 1:nrow(InvestmentScenarios)) {
    PortfolioMaturityModified[i,j] <-
      modifiedSimulation[nrow(modifiedSimulation),j] * InvestmentScenarios[i, "AmountStocks"] +
      PutMaturityModified[4,j] * InvestmentScenarios[i, "AmountPuts"] +
      InvestmentScenarios[i, "Cash"]
  }
}

# mean return from both portfolio values
PortfolioReturnModified = matrix(data = NA, ncol = N, nrow = nrow(InvestmentScenarios))
for(i in 1:nrow(PortfolioReturnModified)){
  PortfolioReturnModified[i,] <- (PortfolioMaturityModified[i,] - rep(10000, N)) / rep(10000, N)
}
PortfolioReturnMatrixModified$PutInvestment <- PortfolioReturnModified[5,]

# risk measures
for(i in 1:nrow(VariableInvestModified)){
  # mean return
VariableInvestModified[i, "mean return"] <- mean(PortfolioReturnModified[i,])
  # volatility
VariableInvestModified[i, "volatility"] <- volatilityFunction(PortfolioReturnModified[i,])
  # skewness
VariableInvestModified[i, "skewness"] <- skewness(PortfolioReturnModified[i,])
  # VaR 95%
VariableInvestModified[i, "VaR 95 %"] <- VaRFunction(PortfolioReturnModified[i,])[1]
  # VaR 99%
VariableInvestModified[i, "VaR 99 %"] <- VaRFunction(PortfolioReturnModified[i,])[2]
}

FinalTable[7,3:7] <- VariableInvestModified[1,2:6]
FinalTable[8,3:7] <- VariableInvestModified[5,2:6]
```

##### variable Strike Price

Strike Price is varied, but Investment in Portfolio is hold constant for 6% of Puts

```{r}
# calculate portfolio value (positions * price) at maturity - different for each N
PortfolioMaturityModified <- matrix(data = NA, ncol = N, nrow = length(StrikePriceSample))
for(j in 1:N){
  for (i in 1:length(StrikePriceSample)) {
    PortfolioMaturityModified[i,j] <-
      modifiedSimulation[nrow(modifiedSimulation),j] * InvestmentScenarios[3, "AmountStocks"] +
      PutMaturityModified[i,j] * InvestmentScenarios[3, "AmountPuts"] +
      InvestmentScenarios[3, "Cash"]
  }
}

# mean return from both portfolio values
PortfolioReturnModified = matrix(data = NA, ncol = N, nrow = length(StrikePriceSample))
for(i in 1:nrow(PortfolioReturnModified)){
  PortfolioReturnModified[i,] <- (PortfolioMaturityModified[i,] - rep(10000, N)) / PortfolioMaturityModified[i,]
}
PortfolioReturnMatrixModified$StrikePrice <- PortfolioReturnModified[2,]

# risk measures
for(i in 1:nrow(VariableStrikeModified)){
  # mean return
VariableStrikeModified[i, "mean return"] <- mean(PortfolioReturnModified[i,])
  # volatility
VariableStrikeModified[i, "volatility"] <- volatilityFunction(PortfolioReturnModified[i,])
  # skewness
VariableStrikeModified[i, "skewness"] <- skewness(PortfolioReturnModified[i,])
  # VaR 95%
VariableStrikeModified[i, "VaR 95 %"] <- VaRFunction(PortfolioReturnModified[i,])[1]
  # VaR 99%
VariableStrikeModified[i, "VaR 99 %"] <- VaRFunction(PortfolioReturnModified[i,])[2]
}

FinalTable[9,3:7] <- VariableStrikeModified[2,2:6]
```

### Visualization

```{r plot_VIII_case1, eval=TRUE}
ggplot(
  data = PortfolioReturnMatrixModified) +
  geom_density(aes(x = withoutPuts,
               color = 'without Puts')) +
  geom_density(aes(x = StrikePrice,
               color = 'with 18 EUR Strike Price')) +
  geom_density(aes(x = PutInvestment,
               color = 'with 12% Puts')) +
  labs(x = "Return in EUR",
      y = "density") +
  ggtitle("Density estimation of Portfolio Return",
          "-20% stock price") +
  xlim(-1,1) +
  theme_bw(base_size = 25) +
  theme(legend.position = c(0.8, 0.8),
        legend.title = element_blank())

# save the graph
ggsave(
  "./figures/PartVIII_case2.png",
  dpi = 320,
  width = 11,
  height = 7
)
```


### Case 3

> A 20 % drop in the stock price and a 5% increase in volatility. 

##### Simulation

At first the simulation has to be modified. For the riskfree Rate the same value is approximated as before in Task VII.
The simulated values are cut in half and resumed from the cut part.

```{r VIII_1}
# create for each N one single simulation and add it back
for(j in 1:N){
  ResultMatrix[,j] <- simulation(
    startPrice = unlist(originalSimulation[nrow(originalSimulation),j] * 0.8),  # 20% percent drop
    riskFreeDaily = mean(dataSim$riskFreeDaily, na.rm = TRUE),
    vola = unlist(volaSimulated[nrow(originalSimulation),j] + 0.05),  # vola increase + 5%
    T0 = nrow(simulatedPaths) - nrow(originalSimulation),
    N = 1
  )
}
modifiedSimulation <- rbind(originalSimulation, ResultMatrix[2:nrow(ResultMatrix),])
```

##### Puts

only prices for maturity have to be calculated newly.

```{r VIII_2}
    # create matrix to store values in
PutMaturityModified <- matrix(data = NA, ncol = N, nrow = length(StrikePriceSample))
    # riskfree Rate is approximated as the mean of all available dates from Part 1
dataSim <- read_csv("./data/dataSim.csv")

    # volatility as log return from Stock Prices
vola1 <- vector(mode = "numeric", length = nrow(modifiedSimulation)-1)
diff <- vector(mode = "numeric", length = nrow(modifiedSimulation)-1)
volaSimulated <- matrix(data = NA, ncol = N, nrow = nrow(modifiedSimulation)-1)
# calculate the volatility for 1 year
for(j in 1:N) {
  for (i in 2:nrow(modifiedSimulation)) {
    diff[i - 1] <-
      log(modifiedSimulation[i - 1, j]) - log(modifiedSimulation[i, j])  # daily stock return
    vola1[i - 1] <-
      sd(diff) * sqrt(traidingDays[1])  # standard deviation of stock return for year 1
  }
  volaSimulated[,j] <- vola1
}

for(i in 1:length(StrikePriceSample)){
  for(j in 1:N){
    PutMaturityModified[i,j] <- BlackScholes(
      StockPrice = unlist(modifiedSimulation[nrow(modifiedSimulation), j]),
      StrikePrice = unlist(StrikePriceSample[i]),
      RiskFreeRate = mean(dataSim$riskFreeDaily, na.rm = TRUE),
      TimeToExpiration = .001,  # approximation to zero
      volatility = unlist(volaSimulated[nrow(volaSimulated),j])
      )[1]
  }
}
  # calculate Gain and Loss for Put Options
for(i in 1:length(StrikePriceSample)){
    # calculate mean price of Put at Maturity 
  PutMatrixModified[i,3] <- mean(PutMaturityModified[,i])
  PutMatrixModified[i,4] <- PutMatrixModified[i,3] - PutMatrixModified[i,2]
}
```


### Portfolio

##### variable Strike Price

```{r VIII_3}
# calculate portfolio value (positions * price) at maturity - different for each N
PortfolioMaturityModified <- matrix(data = NA, ncol = N, nrow = nrow(InvestmentScenarios))
for(j in 1:N){
  for (i in 1:nrow(InvestmentScenarios)) {
    PortfolioMaturityModified[i,j] <-
      modifiedSimulation[nrow(modifiedSimulation),j] * InvestmentScenarios[i, "AmountStocks"] +
      PutMaturityModified[4,j] * InvestmentScenarios[i, "AmountPuts"] +
      InvestmentScenarios[i, "Cash"]
  }
}

# mean return from both portfolio values
PortfolioReturnModified = matrix(data = NA, ncol = N, nrow = nrow(InvestmentScenarios))
for(i in 1:nrow(PortfolioReturnModified)){
  PortfolioReturnModified[i,] <- (PortfolioMaturityModified[i,] - rep(10000, N)) / rep(10000, N)
}
PortfolioReturnMatrixModified$PutInvestment <- PortfolioReturnModified[5,]

# risk measures
for(i in 1:nrow(VariableInvestModified)){
  # mean return
VariableInvestModified[i, "mean return"] <- mean(PortfolioReturnModified[i,])
  # volatility
VariableInvestModified[i, "volatility"] <- volatilityFunction(PortfolioReturnModified[i,])
  # skewness
VariableInvestModified[i, "skewness"] <- skewness(PortfolioReturnModified[i,])
  # VaR 95%
VariableInvestModified[i, "VaR 95 %"] <- VaRFunction(PortfolioReturnModified[i,])[1]
  # VaR 99%
VariableInvestModified[i, "VaR 99 %"] <- VaRFunction(PortfolioReturnModified[i,])[2]
}

FinalTable[10,3:7] <- VariableInvestModified[1,2:6]
FinalTable[11,3:7] <- VariableInvestModified[5,2:6]
```

##### variable Strike Price

Strike Price is varied, but Investment in Portfolio is hold constant for 6% of Puts

```{r}
# calculate portfolio value (positions * price) at maturity - different for each N
PortfolioMaturityModified <- matrix(data = NA, ncol = N, nrow = length(StrikePriceSample))
for(j in 1:N){
  for (i in 1:length(StrikePriceSample)) {
    PortfolioMaturityModified[i,j] <-
      modifiedSimulation[nrow(modifiedSimulation),j] * InvestmentScenarios[3, "AmountStocks"] +
      PutMaturityModified[i,j] * InvestmentScenarios[3, "AmountPuts"] +
      InvestmentScenarios[3, "Cash"]
  }
}

# mean return from both portfolio values
PortfolioReturnModified = matrix(data = NA, ncol = N, nrow = length(StrikePriceSample))
for(i in 1:nrow(PortfolioReturnModified)){
  PortfolioReturnModified[i,] <- (PortfolioMaturityModified[i,] - rep(10000, N)) / PortfolioMaturityModified[i,]
}
PortfolioReturnMatrixModified$StrikePrice <- PortfolioReturnModified[2,]

# risk measures
for(i in 1:nrow(VariableStrikeModified)){
  # mean return
VariableStrikeModified[i, "mean return"] <- mean(PortfolioReturnModified[i,])
  # volatility
VariableStrikeModified[i, "volatility"] <- volatilityFunction(PortfolioReturnModified[i,])
  # skewness
VariableStrikeModified[i, "skewness"] <- skewness(PortfolioReturnModified[i,])
  # VaR 95%
VariableStrikeModified[i, "VaR 95 %"] <- VaRFunction(PortfolioReturnModified[i,])[1]
  # VaR 99%
VariableStrikeModified[i, "VaR 99 %"] <- VaRFunction(PortfolioReturnModified[i,])[2]
}

FinalTable[12,3:7] <- VariableStrikeModified[2,2:6]
```

### Visualization

```{r plot_VIII_case1, eval=TRUE}
ggplot(
  data = PortfolioReturnMatrixModified) +
  geom_density(aes(x = withoutPuts,
               color = 'without Puts')) +
  geom_density(aes(x = StrikePrice,
               color = 'with 18 EUR Strike Price')) +
  geom_density(aes(x = PutInvestment,
               color = 'with 12% Puts')) +
  labs(x = "Return in EUR",
      y = "density") +
  ggtitle("Density estimation of Portfolio Return",
          "-20% stock price & +5% volatility") +
  xlim(-1,1.25) +
  theme_bw(base_size = 25)
  theme(legend.position = c(0.8, 0.8),
        legend.title = element_blank())

# save the graph
ggsave(
  "./figures/PartVIII_case3.png",
  dpi = 320,
  height = 7,
  width = 11
)

FinalTable
```